---
title: "PBN RNAseq DeSeq2 RNA-Sequencing Analysis -- By Sex"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(reshape2)
library(gridExtra)
library(ggrepel)
library(plotly)
library(kableExtra)
library(pheatmap)
library(RColorBrewer)
library(knitr)
library(DESeq2)
library(clusterProfiler)
library(org.Mm.eg.db)
library(vidger)
library(ggrepel)
library(DOSE)
```

# DESeq2 Analysis
DESeq2 is called our post-processing
## Setup and Analysis
###Building DESeq2 input files from gene count matrices output by String Tie.
```{r DESeq2 Build, message=FALSE, warning=FALSE, include=FALSE}

# Establish Path Variable
path="~/Documents/GitHubRepos/CGRPReinforcement/"

# The following file has to be without spaces and needs to be exactly the same as the example sample info
colData <- read.csv(paste0(path, "data/sample_info_PBN_CGRP.csv"), sep = ",", row.names = 1, check.names = FALSE)
# this is the gene count matrix which is the output from the preprocessing shell, the following command does not work from transcript count matrix
countData <-as.matrix(read.csv(paste0(path, "data/gene_count_matrix_LT19_PBN2_LT15_Merged.csv"), row.names = "gene_id", check.names = FALSE))

#colnames(countData)
#rownames(colData)
## Following needs to return true
#all(rownames(colData) %in% colnames(countData))
## if there are any samples in the gene count matrix that should not be in there the following command will remove them
countData <- countData[, rownames(colData)]
#all(rownames(colData) == colnames(countData))
```

###Running DeSeq2 analysis with factor levels of Sex
```{r DESeq2 Analysis, echo=FALSE, message=FALSE, warning=FALSE}
colData$sex <- factor(colData$sex)

dds <- DESeqDataSetFromMatrix(countData = countData, colData = colData, design = ~sex)

## Filter out counts less than 50 across the rows
## This helps to remove noise from the data set
keep <- rowSums(counts(dds)) >= 50
dds <- dds[keep,]

## Re-level Conditions to analyze against HA-
dds$sex <- relevel(dds$sex, ref = "Female")
dds <- DESeq(dds)

#resultsNames(dds)

plotDispEsts(dds)
```

## Plot Estimates and Gather Results

These results include overview for each result file and dispersion estimates based on count values

### Dispersion Estimates Male vs Female based on count values
```{r HA+ vs HA- Results, echo=FALSE, paged.print=TRUE, results='asis'}
res1 <- results(dds, name = "sex_Male_vs_Female")

kable(head(res1, 20), caption = "Male vs Female Results Overview", format = "html", align = "l", booktabs = TRUE) %>%
  kable_styling(html_font = "Arial", font_size = 10, bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(width = "1000px", height = "500px")

summary(res1)
```

## Significant DEGs By Comparison

This subsection is for annotation the results files to include all necessary information for downstream analyses. This is due to the use of StringTie to generate count matrices with the specific mm10 annotation files that we use. Other methods will probably not require this, but our gene names are split by the "|" delimiter and must be separated in order to accurately generate the required Gene Symbols for mapping to entrez, ensemble and descriptions from Org.mm.eg.db. 
```{r Annotate DDS Matrices, message=FALSE, warning=FALSE, include=FALSE, results='asis'}
## Make new rownames file and split based on pipe from dds file
deseqrownames <- data.frame(rownames(dds))
newrownames <- deseqrownames %>% tidyr::separate(rownames.dds., c("id", "symbol"),sep ="([|])")

#Check to make sure they split correctly and do not overlap
all(newrownames$id == newrownames$symbol)

## this should be 0 for everything
sum(is.na(newrownames$symbol))
sum(duplicated(newrownames$symbol))

mcols(dds) <- cbind(mcols(dds), newrownames)

# Add gene symbol to results
res1$symbol <- mcols(dds)$symbol

# Create a list to store the result objects
result_list <- list()

# Define the number of results
num_results <- 1

# Loop through each result
for (i in 1:num_results) {
  # Get the result object name
  result_name <- paste0("res", i)
  
  # Get the DESeq2 result object
  result_obj <- get(result_name)
  
  # Map symbol to gene description
  result_obj$description <- mapIds(x = org.Mm.eg.db,
                                    keys = result_obj$symbol,
                                    column = "GENENAME",
                                    keytype = "SYMBOL",
                                    multiVals = "first")
  
  # Add ENTREZ ID
  result_obj$entrez <- mapIds(x = org.Mm.eg.db,
                               keys = result_obj$symbol,
                               column = "ENTREZID",
                               keytype = "SYMBOL",
                               multiVals = "first")
  
  # Add ENSEMBL ID
  result_obj$ensembl <- mapIds(x = org.Mm.eg.db,
                                keys = result_obj$entrez,
                                column = "ENSEMBL",
                                keytype = "ENTREZID",
                                multiVals = "first")
  
  # Store the result object in the list
  result_list[[i]] <- result_obj

  # Write out the subsetted result object to a variable with name results_sig_i
  assign(paste0("res", i), result_obj)
}
```

After annotation we subset the results to only view the significantly differentially expressed genes with an adjusted p-value \> 0.05.

```{r Subset Significant Results, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='asis'}
results_sig1 <- subset(res1, padj < 0.05)

kable((head(results_sig1, 20)), caption = "Significant DEGs Male vs Female", format = "html", align = "l", booktabs = TRUE) %>%
  kable_styling(html_font = "Arial", font_size = 10, bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(width = "1000px", height = "500px")
```



```{r eval=FALSE, include=FALSE}

#write_csv(as.data.frame(results_sig1),paste0(path, "output/SigDEGs_PBNRNAseq_bySex.csv"))

```

## Visualizing Differentially Expressed Genes

### Principal Components Analysis

For visualization I am converting all analyses to logarithmic. Following this I have completed a PCA based on the top 500 genes.

```{r PCA Plots, echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
#rlog transformation of dds for visualization
ddsMat_rlog <- rlog(dds, blind = FALSE)

# Plot PCA by column variable (dds)

pcaData2 <- plotPCA(ddsMat_rlog, intgroup = "sex", returnData = TRUE)
pcaData2$sex <- factor(pcaData2$sex, levels = c("Male", "Female"))
percentVar2 <- round(100 * attr(pcaData2, "percentVar"))
## you only need the following line but the rest is additional formatting
pca <- ggplot(pcaData2, aes(PC1, PC2, color=sex, label = colnames(dds))) +
  theme_bw() + # remove default ggplot2 theme
  geom_point(size = 2) + # Increase point size
  scale_y_continuous(limits = c(-125, 125)) +# change limits to fix figure dimensions
  scale_x_continuous(limits = c(-125, 125)) +# change limits to fix figure dimensions
  xlab(paste0("PC1: ",percentVar2[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar2[2],"% variance")) + 
  ggrepel::geom_label_repel(box.padding = 0.5, max.overlaps = Inf) +
  coord_fixed() +
  #scale_color_manual(values = custom_colors) +  # Set custom colors
  ggtitle(label = "Supp Figure 2A. Principal Component Analysis (PCA)", 
          subtitle = "Top 500 Variable Genes (rlog transformed)")

pca

#pdf(paste0(path, "output/SFigure2A_PCA_bySex.pdf"), width = 10, height = 8)
#pca
#dev.off()
```

### Heatmaps of Significantly Differentially Expressed Genes {.tabset}

To generate the heatmaps we use two methods, the first utilizes all differential expressed genes based on a specified cut-off. In our case we select for any significant DEGs with a L2FC \> 1.5 or an L2FC \< -1.5 with a L2FC Standard Error \< 0.75. We then plot these using a wards D correlation clustering method and generate both supervised and unsupervised clusters (by group). Rows are scaled based on a correlation z-score. The second method is less stringent and utilizes all significant DEGs, without filtering.

#### Highly Significiant

Cut-Off of L2FC \> 1.5 or an L2FC \< -1.5 with a L2FC Standard Error \< 0.75

```{r include=FALSE}
ressig1 <- as.data.frame(results_sig1)

ressig1$Comparison <- "Male vs Female"

ressig1_highlfc_pos <- ressig1[ressig1$log2FoldChange >= 1, ]
ressig1_highlfc_neg <- ressig1[ressig1$log2FoldChange <= -1, ]

temp_sigdegshm <- rbind(ressig1_highlfc_pos, ressig1_highlfc_neg)

## the following step limits the error allowed on each gene, which will make all genes look more similar, there is a lot of variablity in this dataset, cannot be set to less than 1
temp_sigdegshm1 <- temp_sigdegshm[temp_sigdegshm$lfcSE <=1.5, ]

dup_rows <- duplicated(temp_sigdegshm1$symbol)
data_dedup <- temp_sigdegshm1[!dup_rows, ]

rows <- rownames(data_dedup)

select <- order(rowMeans(counts(dds, normalized = TRUE)),
decreasing = TRUE)

mat <- assay(ddsMat_rlog)[select, ]

data_mod <- mat[rownames(mat) %in% rows, ]

matrownames_data_mod <- data.frame(rownames(data_mod))
matrownames_data_mod <- matrownames_data_mod %>% tidyr::separate(rownames.data_mod., c("id", "symbol"), sep = "([|])")

samples <- tibble::as_tibble(read.csv(paste0(path, "data/sample_info_PBN_CGRP.csv"), sep = ","))

annot_col <- samples %>%
tibble::column_to_rownames("sampleName") %>%
as.data.frame()

custom_colors <- c("Male" = "green",
                   "Female" = "purple")

ann_colors <- list(
condition = as.factor(custom_colors),
group = as.factor(custom_colors))


```

##### Supervised and Condensed Highly Significant DEGs (L2FC > +/- 1 and L2FCSE < 1.5).

```{r Sup HM Highly Sig DEGs, echo=FALSE, fig.height=8, fig.width=8}

pheatmap::pheatmap(
mat = data_mod,
color = colorRampPalette(rev(RColorBrewer::brewer.pal(11, "RdYlBu")))(255), ## this just changes the colors
scale = "row", ## based on Rlog fold change
annotation_colors = ann_colors,  ## assigns group colors
annotation_col = annot_col,  ## assigns group names
fontsize = 6,
show_colnames = T,
cluster_rows = T, ## puts the lows together and the highs together
cluster_cols = F,  ## when this is T this is unsupervised
annotation_legend = T, ## this is legend for the treatment conditions
legend = T, ## legend for color
show_rownames = F, ## this is gene names but with over 2,000 this isn't useful
clustering_distance_rows = "correlation",
clustering_method = "ward.D",
main = "Heatmap of Significant DEGs (L2FC > +/- 1.50 and L2FCSE < 0.75)-- Supervised and Condensed", ## Title
cluster_row_slices = TRUE,
show_row_dend = FALSE,
gaps_col = c(4),
cutree_rows = 8,  ## this separates the genes into 6 blocks
cutree_cols = 2, ## this separates the conditions into 4 blocks
treeheight_row = 15,  ## this will only show up if clustering by row
treeheight_col = 10)  ## this will only show up if clustering by column
#filename = paste0("output/PBNRNASeq_Heatmap_HSDegs_Supervised_Condensed.pdf") ## Only include if saving to a file

```

##### Unsupervised and Condensed Highly Significant DEGs (L2FC > +/- 1 and L2FCSE < 1.5)

```{r echo=FALSE, fig.height=8, fig.width=8}
pheatmap::pheatmap(
mat = data_mod,
color = colorRampPalette(rev(RColorBrewer::brewer.pal(11, "RdYlBu")))(250),
scale = "row",
annotation_colors = ann_colors,
annotation_col = annot_col,
fontsize = 6,
show_colnames = T,
cluster_rows = T,
cluster_cols = T,
annotation_legend = T,
legend = T,
show_rownames = F,
clustering_distance_rows = "correlation",
clustering_method = "ward.D",
main = "Heatmap of Significant DEGs (L2FC > +/- 1.50 and L2FCSE < 0.75) -- Unsupervised and Condensed",
cluster_row_slices = TRUE,
show_row_dend = FALSE,
gaps_col = c(3, 5, 8),
cutree_rows = 6,
cutree_cols = 4,
treeheight_row = 15,
treeheight_col = 10)
#filename = paste0("output/PBNRNASeq_Heatmap_HSDegs_Unsupervised_Condensed.pdf")
```

#### All DEGs

```{r include=FALSE}
ressig1 <- as.data.frame(results_sig1)

ressig1$Comparison <- "Male vs Female"

dup_rows <- duplicated(ressig1$symbol)
data_dedup <- ressig1[!dup_rows, ]

rows <- rownames(data_dedup)

select <- order(rowMeans(counts(dds, normalized = TRUE)),
decreasing = TRUE)

mat <- assay(ddsMat_rlog)[select, ]

data_mod <- mat[rownames(mat) %in% rows, ]

matrownames_data_mod <- data.frame(rownames(data_mod))
matrownames_data_mod <- matrownames_data_mod %>% tidyr::separate(rownames.data_mod., c("id", "symbol"), sep = "([|])")


```

##### Supervised and Condensed All DEGs

```{r echo=FALSE, fig.height=8, fig.width=8}
pheatmap::pheatmap(
mat = data_mod,
color = colorRampPalette(rev(RColorBrewer::brewer.pal(11, "RdYlBu")))(255),
scale = "row",
annotation_colors = ann_colors,
annotation_col = annot_col,
fontsize = 6,
show_colnames = T,
cluster_rows = T,
cluster_cols = F,
annotation_legend = T,
legend = T,
show_rownames = F,
clustering_distance_rows = "correlation",
clustering_method = "ward.D",
main = "Heatmap of Significant DEGs -- Supervised and Condensed",
cluster_row_slices = TRUE,
show_row_dend = FALSE,
gaps_col = 4,
cutree_rows = 2,
cutree_cols = 4,
treeheight_row = 15,
treeheight_col = 10)
#filename = paste0("output/PBNRNASeq_Heatmap_AllDegs_Supervised_Condensed.pdf")) 
```

##### Unsupervised and Condensed All DEGs

```{r echo=FALSE, fig.height=8, fig.width=8}
pheatmap::pheatmap(
mat = data_mod,
color = colorRampPalette(rev(RColorBrewer::brewer.pal(11, "RdYlBu")))(250),
scale = "row",
annotation_colors = ann_colors,
annotation_col = annot_col,
fontsize = 6,
show_colnames = T,
cluster_rows = T,
cluster_cols = T,
annotation_legend = T,
legend = T,
show_rownames = F,
clustering_distance_rows = "correlation",
clustering_method = "ward.D",
main = "Supp Fig 2B. Heatmap of Significant DEGs -- Unsupervised and Condensed",
cluster_row_slices = TRUE,
show_row_dend = FALSE,
gaps_col = c(4,8,12,16),
cutree_rows = 4,
cutree_cols = 4,
treeheight_row = 15,
treeheight_col = 10)
#filename = paste0(path, "output/SFigure2B_Heatmap_Unsupervised_Condensed.pdf")) 
```

# Gene Ontology Analysis {.tabset}
To generate GO analysis we first created a background of genes using any genes that met the original DESEQ2 filtering criteria of \> 50 genes counts per row and having a mappable Entrez ID. This generates the necessary background to ensure that we are not just detecting background functional pathways as we are utilizing this specific cell type. We then filter our genes that have L2FC \< 0 or L2FC \> 0 to determine directionality of the affected GO pathways. GO enrich is then used to conduct the pathway analysis utilizing the Entrez IDs as inputs and an "FDR" multiple testing correction.

### Pull only the Enriched samples from either side  (<-1 or >1 LFC)
```{r}
# Generate Background for KEGG/GO
background_mm10 <- subset(res1, is.na(entrez) == FALSE)
gene_matrix_background <- unique(sort(background_mm10$entrez))

# Remove any genes that do not have any entrez identifiers
results_sig_entrez <- subset(results_sig1, is.na(entrez) == FALSE)

# Create a matrix of gene log2 fold changes
gene_matrix <- results_sig_entrez$log2FoldChange

# Add the entrezID's as names for each logFC entry
names(gene_matrix) <- results_sig_entrez$entrez

# Pull only the Enriched samples from either side +/- 1 LFC
gene_matrix_enrich <- gene_matrix[which(gene_matrix > 1)]
gene_matrix_enrich2 <- gene_matrix[which(gene_matrix < -1)]

go_enrich1 <- clusterProfiler::enrichGO(gene = names(gene_matrix_enrich),
                      OrgDb = 'org.Mm.eg.db', 
                      readable = T,
                      universe = gene_matrix_background,
                      ont = "ALL",
                      pAdjustMethod = "fdr",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.05)

go_enrich2 <- clusterProfiler::enrichGO(gene = names(gene_matrix_enrich2),
                      OrgDb = 'org.Mm.eg.db', 
                      readable = T,
                      universe = gene_matrix_background,
                      ont = "ALL",
                      pAdjustMethod = "fdr",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.05)

go_enrich_dataset1 <- as.data.frame(go_enrich1)
go_enrich_dataset2 <- as.data.frame(go_enrich2)

go_enrich_dataset1$Dataset <- "Male vs Female"
go_enrich_dataset2$Dataset <- "Female vs Male"

go_enrich_dataset1$Celltype <- "Male"
go_enrich_dataset2$Celltype <- "Female"

go_enrich_dataset1$Comparison <- "Female"
go_enrich_dataset2$Comparison <- "Male"

MergedGO1 <- merge(go_enrich_dataset1, go_enrich_dataset2, all=TRUE)

# Turn your 'Dataset' column into a character vector
MergedGO1$Dataset <- as.character(MergedGO1$Dataset)
# Then turn it back into a factor with the levels in the correct order
MergedGO1$Dataset <- factor(MergedGO1$Dataset, levels = c("Male vs Female", "Female vs Male"))

MergedGO1$Description_title_case <- stringr::str_to_title(MergedGO1$Description)
MergedGO1$GOPathways <- stringr::str_to_title(MergedGO1$Description_title_case)
MergedGO1$GOPathways <- stringr::str_wrap(MergedGO1$GOPathways, width = 10)

## write results from GO1 to a csv
write.csv(as.data.frame(MergedGO1),paste0(path, "output/PBNRNASeq_bySex_GOFullAnalysis.csv"))

S1 <- ggplot(MergedGO1, aes(x = Dataset, y = Description, size = Count, color = qvalue, group = Dataset)) +
  geom_point(alpha = 1) +
  theme_linedraw() + # change legend key width
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15)) +
  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 50)) +
  ggtitle("Enriched Pathways in PBN Nuclei Cell Types")
  #facet_grid(~Region+Comparison2, scales = "free_x")

S2 <- S1 + scale_color_gradient(low = "red2", high = "mediumblue", space = "Lab") + scale_size(range = c(5, 10))
S2
```
### Set the Fold Enrichment for most important pathways

```{r}
go_filtered1 <- go_enrich_dataset1 %>%
  filter(FoldEnrichment >= 2)

go_filtered2 <- go_enrich_dataset2 %>%
  filter(FoldEnrichment >= 2)

go_filtered1$Dataset <- "Male vs Female"
go_filtered2$Dataset <- "Female vs Male"

go_filtered1$Celltype <- "Male"
go_filtered2$Celltype <- "Female"

go_filtered1$Comparison <- "Female"
go_filtered2$Comparison <- "Male"

```

#### Combined Filtered
```{r}
MergedGO1_filtered <- merge(go_filtered1, go_filtered2, all=TRUE)

# Turn your 'Dataset' column into a character vector
MergedGO1_filtered$Dataset <- as.character(MergedGO1_filtered$Dataset)
# Then turn it back into a factor with the levels in the correct order
MergedGO1_filtered$Dataset <- factor(MergedGO1_filtered$Dataset, levels = c("Male vs Female", "Female vs Male"))

MergedGO1_filtered$Description_title_case <- stringr::str_to_title(MergedGO1_filtered$Description)
MergedGO1_filtered$GOPathways <- stringr::str_to_title(MergedGO1_filtered$Description_title_case)
MergedGO1_filtered$GOPathways <- stringr::str_wrap(MergedGO1_filtered$GOPathways, width = 10)

S1 <- ggplot(MergedGO1_filtered, aes(x = Dataset, y = Description, size = Count, color = qvalue, group = Dataset)) +
  geom_point(alpha = 1) +
  theme_linedraw() + # change legend key width
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15)) +
  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 50)) +
  ggtitle("Enriched Pathways in PBN Nuclei Cell Types")
  #facet_grid(~Region+Comparison2, scales = "free_x")

S2 <- S1 + scale_color_gradient(low = "red2", high = "mediumblue", space = "Lab") + scale_size(range = c(5, 10))
S2

```

#### Male Only

```{r}
GO1_filtered <- go_filtered1

# Turn your 'Dataset' column into a character vector
GO1_filtered$Dataset <- as.character(GO1_filtered$Dataset)
# Then turn it back into a factor with the levels in the correct order
GO1_filtered$Dataset <- factor(GO1_filtered$Dataset, levels = c("Male vs Female"))

GO1_filtered$Description_title_case <- stringr::str_to_title(GO1_filtered$Description)
GO1_filtered$GOPathways <- stringr::str_to_title(GO1_filtered$Description_title_case)
GO1_filtered$GOPathways <- stringr::str_wrap(GO1_filtered$GOPathways, width = 10)

S1 <- ggplot(GO1_filtered, aes(x = Count, y = reorder(Description, Count), fill = qvalue)) +
  geom_col(alpha = 1) +
  facet_wrap(~Dataset, scales = "free_y") +
  theme_linedraw() + # change legend key width
  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 50)) +
  ggtitle("Supp Figure 2C: Enriched Pathways in Male CGRPPBN Neurons")

S2 <- S1 + scale_color_gradient(low = "red2", high = "mediumblue", space = "Lab") + scale_size(range = c(5, 10))
S2

pdf(paste0(path, "output/SFig2D_GO_SexSpecific.pdf"), width = 10, height = 8)
S2
dev.off()
```

#### Female Only

```{r}
GO2_filtered <- go_filtered2

# Turn your 'Dataset' column into a character vector
GO2_filtered$Dataset <- as.character(GO2_filtered$Dataset)
# Then turn it back into a factor with the levels in the correct order
GO2_filtered$Dataset <- factor(GO2_filtered$Dataset, levels = c("Female vs Male"))

GO2_filtered$Description_title_case <- stringr::str_to_title(GO2_filtered$Description)
GO2_filtered$GOPathways <- stringr::str_to_title(GO2_filtered$Description_title_case)
GO2_filtered$GOPathways <- stringr::str_wrap(GO2_filtered$GOPathways, width = 10)

S1 <- ggplot(GO2_filtered, aes(x = Count, y = reorder(Description, Count), fill = qvalue)) +
  geom_col(alpha = 1) +
  facet_wrap(~Dataset, scales = "free_y") +
  theme_linedraw() + # change legend key width
  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 50)) +
  ggtitle("Supp Figure 2C: Enriched Pathways in Female CGRPPBN Neurons")

S2 <- S1 + scale_color_gradient(low = "red2", high = "mediumblue", space = "Lab") + scale_size(range = c(5, 10))
S2


#pdf(paste0(path, "output/SFig2E_GO_SexSpecific.pdf"), width = 10, height = 8)
#S2
#dev.off()
```


```{r}

sessionInfo()

```
