---
title: "PBN RNAseq DeSeq2 RNA-Sequencing Analysis -- By HA"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(reshape2)
library(gridExtra)
library(ggrepel)
library(plotly)
library(kableExtra)
library(pheatmap)
library(RColorBrewer)
library(knitr)
library(DESeq2)
library(clusterProfiler)
library(org.Mm.eg.db)
library(vidger)
library(ggrepel)
library(DOSE)
```

# Experiment Information

<center><font size="18"> **Parabrachial CGRP Neurons Regulate Opioid
Reinforcement** </font></center>

<center>Lauren L. Bystrom1,2,4, Alexander V. Margetts1-3, Nicole M.
Kujas1-3, Florence M Bourgain-Guglielmetti1-3, Luis M.
Tuesta1-3\*</center>

<center>1Department of Psychiatry and Behavioral Sciences 2Center for
Therapeutic Innovation 3Sylvester Comprehensive Cancer Center 4Medical
Scientist Training Program University of Miami Miller School of
Medicine, Miami, FL 33136 \* Corresponding author
([ltuesta\@miami.edu](mailto:ltuesta@miami.edu){.email})</center>

# Abstract

Opioid use disorder (OUD) is a chronic, relapsing disease driven by the
reinforcing properties of opioids and perpetuated by avoidance of the
negative affective states associated with withdrawal. While available
OUD pharmacotherapies can be effective, there is an urgent need to
identify novel therapeutic targets. One potential target is calcitonin
gene-related peptide (CGRP). CGRP is a neuropeptide produced by a small
population of neurons in the parabrachial nucleus (CGRPPBN). In this
study, we use a virally mediated cell-specific isolation approach in
combination with nuclear RNA-sequencing to establish the baseline
transcriptome of CGRPPBN neurons in mice. Multiple genes related to
appetite were highly enriched in CGRPPBN neurons, consistent with
previous literature on the behavioral role of these neurons.
Interestingly, CGRPPBN neurons also highly express the µ-opioid
receptor, suggesting these neurons may be involved in opioid
reinforcement. To this end, we show that CGRPPBN neurons are sensitive
to morphine administration and withdrawal based on cFos staining,
suggesting that this population may play a role in regulating the
appetitive aspects of opioid taking. To test this hypothesis, we
employed a dose-response model of morphine intravenous
self-administration (IVSA) in mice, in which DREADD-mediated inhibition
of CGRPPBN neurons decreased morphine intake at multiple morphine doses,
but did not affect morphine tolerance or context-induced morphine
seeking following prolonged abstinence. These data suggest CGRPPBN
neurons play an important role in morphine reinforcement, possibly via
appetitive control, and may therefore serve as a novel therapeutic
target for OUD.

# DESeq2 Analysis

DESeq2 is used for Differential Gene Expression Analyses

## Data Setup
Building DESeq2 input files from gene count matrices output by String Tie.

```{r DESeq2 Build, message=FALSE, warning=FALSE, include=FALSE}
## Run DEeq2 from Matrix output using Stringtie to compare all groups
#setwd("~/Documents/Raw_Sequencing_Data/GitHubRepos/CGRPReinforcement/")

# Establish Path Variable
path="~/Documents/GitHubRepos/CGRPReinforcement/"

# The following file has to be without spaces and needs to be exactly the same as the example sample info
colData <- read.csv(paste0(path, "data/sample_info_PBN_DG_Exclusions.csv"), sep = ",", row.names = 1, check.names = FALSE)

# This is the gene count matrix which is the output from the preprocessing shell, the following command does not work from transcript count matrix
countData <-as.matrix(read.csv(paste0(path, "data/gene_count_matrix_LT19_PBN2_LT15_Merged.csv"), row.names = "gene_id", check.names = FALSE))

#colnames(countData)
#rownames(colData)
## Following needs to return true
#all(rownames(colData) %in% colnames(countData))
## if there are any samples in the gene count matrix that should not be in there the following command will remove them
countData <- countData[, rownames(colData)]
#all(rownames(colData) == colnames(countData))

```

## Running DeSeq2 analysis with factor levels of HA tag

```{r DESeq2 Analysis, echo=TRUE, fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
colData$Hatag <- factor(colData$Hatag)

dds <- DESeqDataSetFromMatrix(countData = countData, colData = colData, design = ~Hatag)

## Filter out counts less than 50 across the rows
## This helps to remove noise from the data set
keep <- rowSums(counts(dds)) >= 50
dds <- dds[keep,]

## Re-level Conditions to analyze against HA-
dds$Hatag <- relevel(dds$Hatag, ref = "negative")
dds <- DESeq(dds)

#resultsNames(dds)

plotDispEsts(dds)
```

## Plot Estimates and Gather Results

These results include overview for each result file and dispersion
estimates based on count values

### Dispersion Estimates HA+ vs HA- based on count values

```{r HA+ vs HA- Results, echo=FALSE, results='asis'}
res1 <- results(dds, name = "Hatag_positive_vs_negative")

kable(head(res1, 20), caption = "HA+ vs HA- Results Overview", format = "html", align = "l", booktabs = TRUE) %>%
  kable_styling(html_font = "Arial", font_size = 10, bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(width = "1000px", height = "500px")

summary(res1)
```

## Significant DEGs By Comparison

This subsection is for annotation the results files to include all
necessary information for downstream analyses. This is due to the use of
StringTie to generate count matrices with the specific mm10 annotation
files that we use. Other methods will probably not require this, but our
gene names are split by the "\|" delimiter and must be separated in
order to accurately generate the required Gene Symbols for mapping to
entrez, ensemble and descriptions from Org.mm.eg.db.

```{r Annotate DDS Matrices, message=FALSE, warning=FALSE, include=FALSE, results='asis'}
## Make new rownames file and split based on pipe from dds file
deseqrownames <- data.frame(rownames(dds))
newrownames <- deseqrownames %>% tidyr::separate(rownames.dds., c("id", "symbol"),sep ="([|])")

#Check to make sure they split correctly and do not overlap
all(newrownames$id == newrownames$symbol)

## this should be 0 for everything
sum(is.na(newrownames$symbol))
sum(duplicated(newrownames$symbol))

mcols(dds) <- cbind(mcols(dds), newrownames)

# Add gene symbol to results
res1$symbol <- mcols(dds)$symbol

# Create a list to store the result objects
result_list <- list()

# Define the number of results
num_results <- 1

# Loop through each result
for (i in 1:num_results) {
  # Get the result object name
  result_name <- paste0("res", i)
  
  # Get the DESeq2 result object
  result_obj <- get(result_name)
  
  # Map symbol to gene description
  result_obj$description <- mapIds(x = org.Mm.eg.db,
                                    keys = result_obj$symbol,
                                    column = "GENENAME",
                                    keytype = "SYMBOL",
                                    multiVals = "first")
  
  # Add ENTREZ ID
  result_obj$entrez <- mapIds(x = org.Mm.eg.db,
                               keys = result_obj$symbol,
                               column = "ENTREZID",
                               keytype = "SYMBOL",
                               multiVals = "first")
  
  # Add ENSEMBL ID
  result_obj$ensembl <- mapIds(x = org.Mm.eg.db,
                                keys = result_obj$entrez,
                                column = "ENSEMBL",
                                keytype = "ENTREZID",
                                multiVals = "first")
  
  # Store the result object in the list
  result_list[[i]] <- result_obj

  # Write out the subsetted result object to a variable with name results_sig_i
  assign(paste0("res", i), result_obj)
}
```

After annotation we subset the results to only view the significantly
differentially expressed genes with an adjusted p-value \> 0.05.

```{r Subset Significant Results, echo=TRUE, message=FALSE, warning=FALSE, results='asis'}
results_sig1 <- subset(res1, padj < 0.05)

kable((head(results_sig1, 20)), caption = "Significant DEGs HA+ vs HA-", format = "html", align = "l", booktabs = TRUE) %>%
  kable_styling(html_font = "Arial", font_size = 10, bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(width = "1000px", height = "500px")
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

readr::write_csv(as.data.frame(results_sig1), paste0(path, "output/SigDEGs_PBNRNAseq_byHATag.csv"))

```

## Visualizing Differentially Expressed Genes

### Principal Components Analysis (Figure 1B)

For visualization I am converting all analyses to logarithmic. Following
this I have completed a PCA based on the top 500 genes.

```{r PCA Plots, echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
#rlog transformation of dds for visualization
ddsMat_rlog <- rlog(dds, blind = FALSE)

# Plot PCA by column variable (dds)

pcaData2 <- plotPCA(ddsMat_rlog, intgroup = "Hatag", returnData = TRUE)
pcaData2$Hatag <- factor(pcaData2$Hatag, levels = c("negative", "positive"))
percentVar2 <- round(100 * attr(pcaData2, "percentVar"))
## you only need the following line but the rest is additional formatting
pca <- ggplot(pcaData2, aes(PC1, PC2, color=Hatag, label = colnames(dds))) +
  theme_bw() + # remove default ggplot2 theme
  geom_point(size = 2) + # Increase point size
  scale_y_continuous(limits = c(-125, 125)) +# change limits to fix figure dimensions
  scale_x_continuous(limits = c(-125, 125)) +# change limits to fix figure dimensions
  xlab(paste0("PC1: ",percentVar2[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar2[2],"% variance")) + 
  ggrepel::geom_label_repel(box.padding = 0.5, max.overlaps = Inf) +
  coord_fixed() +
  #scale_color_manual(values = custom_colors) +  # Set custom colors
  ggtitle(label = "Figure 1B. Principal Component Analysis (PCA)", 
          subtitle = "Top 500 Most Variable Genes (rlog transformed)")

pca


pdf(paste0(path,"output/Figure1B_PBNRNAseq_PCA_byHATag.pdf"), width = 10, height = 8)
pca
dev.off()

```

### Heatmaps of Significantly Differentially Expressed Genes {.tabset}

To generate the heatmaps we use two methods, the first utilizes all
differential expressed genes based on a specified cut-off. In our case
we select for any significant DEGs with a L2FC \> 1.5 or an L2FC \< -1.5
with a L2FC Standard Error \< 0.75. We then plot these using a wards D
correlation clustering method and generate both supervised and
unsupervised clusters (by group). Rows are scaled based on a correlation
z-score. The second method is less stringent and utilizes all
significant DEGs, without filtering.

#### Highly Significiant

Cut-Off of L2FC \> 1.5 or an L2FC \< -1.5 with a L2FC Standard Error \<
0.75

```{r include=FALSE}
ressig1 <- as.data.frame(results_sig1)

ressig1$Comparison <- "negative vs positive"

ressig1_highlfc_pos <- ressig1[ressig1$log2FoldChange >= 1.50, ]
ressig1_highlfc_neg <- ressig1[ressig1$log2FoldChange <= -1.50, ]

temp_sigdegshm <- rbind(ressig1_highlfc_pos, ressig1_highlfc_neg)

## the following step limits the error allowed on each gene, which will make all genes look more similar
temp_sigdegshm1 <- temp_sigdegshm[temp_sigdegshm$lfcSE <=0.75, ]

dup_rows <- duplicated(temp_sigdegshm1$symbol)
data_dedup <- temp_sigdegshm1[!dup_rows, ]

rows <- rownames(data_dedup)

select <- order(rowMeans(counts(dds, normalized = TRUE)),
decreasing = TRUE)

mat <- assay(ddsMat_rlog)[select, ]

data_mod <- mat[rownames(mat) %in% rows, ]

matrownames_data_mod <- data.frame(rownames(data_mod))
matrownames_data_mod <- matrownames_data_mod %>% tidyr::separate(rownames.data_mod., c("id", "symbol"), sep = "([|])")

samples <- tibble::as_tibble(read.csv(paste0(path, "data/sample_info_PBN_DG_Exclusions.csv"), sep = ","))

custom_colors <- c("positive" = "green", "negative" = "#00CBFF")

annot_col <- samples %>%
tibble::column_to_rownames("sampleName") %>%
as.data.frame()

ann_colors <- list(
condition = as.factor(custom_colors),
group = as.factor(custom_colors))

length(rownames(data_mod))

```

##### Supervised and Condensed Highly Significant DEGs (L2FC \> +/- 1.50 and L2FCSE \< 0.75).

```{r Sup HM Highly Sig DEGs, echo=FALSE, fig.height=8, fig.width=8}

pheatmap::pheatmap(
mat = data_mod,
color = colorRampPalette(rev(RColorBrewer::brewer.pal(11, "RdYlBu")))(255), ## this just changes the colors
scale = "row", ## based on Rlog fold change
annotation_colors = ann_colors,  ## assigns group colors
annotation_col = annot_col,  ## assigns group names
fontsize = 6,
show_colnames = T,
cluster_rows = T, ## puts the lows together and the highs together
cluster_cols = F,  ## when this is T this is unsupervised
annotation_legend = T, ## this is legend for the treatment conditions
legend = T, ## legend for color
show_rownames = F, ## this is gene names but with over 2,000 this isn't useful
clustering_distance_rows = "correlation",
clustering_method = "ward.D",
main = "Heatmap of Significant DEGs (L2FC > +/- 1.50 and L2FCSE < 0.75)-- Supervised and Condensed", ## Title
cluster_row_slices = TRUE,
show_row_dend = FALSE,
gaps_col = c(8),
cutree_rows = 8,  ## this separates the genes into 6 blocks
cutree_cols = 4, ## this separates the conditions into 4 blocks
treeheight_row = 15,  ## this will only show up if clustering by row
treeheight_col = 10)  ## this will only show up if clustering by column
#filename = paste0("output/PBNRNASeq_Heatmap_HSDegs_Supervised_Condensed.pdf") ## Only include if saving to a file

```

##### Unsupervised and Condensed Highly Significant DEGs (L2FC \> +/- 1.50 and L2FCSE \< 0.75)

```{r echo=FALSE, fig.height=8, fig.width=8}
pheatmap::pheatmap(
mat = data_mod,
color = colorRampPalette(rev(RColorBrewer::brewer.pal(11, "RdYlBu")))(250),
scale = "row",
annotation_colors = ann_colors,
annotation_col = annot_col,
fontsize = 6,
show_colnames = T,
cluster_rows = T,
cluster_cols = T,
annotation_legend = T,
legend = T,
show_rownames = F,
clustering_distance_rows = "correlation",
clustering_method = "ward.D",
main = "Heatmap of Significant DEGs (L2FC > +/- 1.50 and L2FCSE < 0.75) -- Unsupervised and Condensed",
cluster_row_slices = TRUE,
show_row_dend = FALSE,
gaps_col = c(3, 5, 8),
cutree_rows = 6,
cutree_cols = 4,
treeheight_row = 15,
treeheight_col = 10)
#filename = paste0("output/PBNRNASeq_Heatmap_HSDegs_Unsupervised_Condensed.pdf")
```

#### All DEGs

```{r include=FALSE}
ressig1 <- as.data.frame(results_sig1)

ressig1$Comparison <- "negative vs positive"

dup_rows <- duplicated(ressig1$symbol)
data_dedup <- ressig1[!dup_rows, ]

rows <- rownames(data_dedup)

select <- order(rowMeans(counts(dds, normalized = TRUE)),
decreasing = TRUE)

mat <- assay(ddsMat_rlog)[select, ]

data_mod <- mat[rownames(mat) %in% rows, ]

matrownames_data_mod <- data.frame(rownames(data_mod))
matrownames_data_mod <- matrownames_data_mod %>% tidyr::separate(rownames.data_mod., c("id", "symbol"), sep = "([|])")

length(rownames(data_mod))

```

##### Supervised and Condensed All DEGs

```{r echo=FALSE, fig.height=8, fig.width=8}
pheatmap::pheatmap(
mat = data_mod,
color = colorRampPalette(rev(RColorBrewer::brewer.pal(11, "RdYlBu")))(255),
scale = "row",
annotation_colors = ann_colors,
annotation_col = annot_col,
fontsize = 6,
show_colnames = T,
cluster_rows = T,
cluster_cols = F,
annotation_legend = T,
legend = T,
show_rownames = F,
clustering_distance_rows = "correlation",
clustering_method = "ward.D",
main = "Heatmap of Significant DEGs -- Supervised and Condensed",
cluster_row_slices = TRUE,
show_row_dend = FALSE,
gaps_col = 8,
cutree_rows = 2,
cutree_cols = 4,
treeheight_row = 15,
treeheight_col = 10)
#filename = paste0("output/PBNRNASeq_Heatmap_AllDegs_Supervised_Condensed.pdf")) 
```

##### Unsupervised and Condensed All DEGs

```{r echo=FALSE, fig.height=8, fig.width=8}
pheatmap::pheatmap(
mat = data_mod,
color = colorRampPalette(rev(RColorBrewer::brewer.pal(11, "RdYlBu")))(250),
scale = "row",
annotation_colors = ann_colors,
annotation_col = annot_col,
fontsize = 6,
show_colnames = T,
cluster_rows = T,
cluster_cols = T,
annotation_legend = T,
legend = T,
show_rownames = F,
clustering_distance_rows = "correlation",
clustering_method = "ward.D",
main = "Figure 1C. Heatmap of Significant DEGs -- Unsupervised and Condensed",
cluster_row_slices = TRUE,
show_row_dend = FALSE,
gaps_col = c(4,8,12,16),
cutree_rows = 6,
cutree_cols = 4,
treeheight_row = 15,
treeheight_col = 10)
#filename = paste0(path, "output/Figure1C_Heatmap.pdf")) 
```

### Volcano Plots {.tabset}

#### Volcano Plots (noninteractive)

```{r Volcano Plots 1, echo=TRUE}
# Gather Log-fold change and FDR-corrected pvalues from DESeq2 results

# Change pvalues to -log10 (1.3 = 0.05)
data <- data.frame(
  gene = (res1$symbol),
  pval = -log10(res1$padj),
  lfc = res1$log2FoldChange
)

# Remove any rows that have NA as an entry
data <- na.omit(data)

# Color the points which are up or down
## If fold-change > 0 and pvalue > 1.3 (Increased significant)
## If fold-change < 0 and pvalue > 1.3 (Decreased significant)

data <- dplyr::mutate(data, color = case_when(
  data$lfc > 0 & data$pval > 1.3 ~ "Increased in HA+",
  data$lfc < 0 & data$pval > 1.3 ~ "Decreased in HA+",
  data$pval < 1.3 ~ "Nonsignificant"
))

data$genelabels <- ifelse(data$gene == "Adamts4"
                          | data$gene == "Fa2h"
                          | data$gene == "Tmem63a"
                          | data$gene == "Klk6"
                          | data$gene == "Gal"
                          | data$gene == "Chodl"
                          | data$gene == "Calca"
                          | data$gene == "Isl1"
                          | data$gene == "Uts2b"
                          | data$gene == "Glul"
                          | data$gene == "Slc17a6"
                          | data$gene == "Scg2", T, F)

# Make a basic ggplot2 object with x-y values
vol <- ggplot(data, aes(x = lfc, y = pval, color = color))

# Add ggplot2 layers
y.axis.text <- element_text(face = "bold", color = "black", size = 7)
x.axis.text <- element_text(face = "bold", color = "black", size = 7)
plot.title.text <- element_text(face = "bold", color = "black", size = 9, hjust = 0.5)
legend.text <- element_text(face = "bold", color = "black", size = 5)
legend.title.text <- element_text(face = "bold", color = "black", size = 6)
axis.title <- element_text(face = "bold", color = "black", size = 6)
```

```{r Volcano Plots 2, echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
vol_final <- vol +
  ggtitle(label = "Figure 1D. HA+ vs HA-") +
  geom_point(size = 1.5, alpha = 0.8, na.rm = T) +
  scale_color_manual(
    name = "Directionality",
    values = c("Increased in HA+" = "#008B00", "Decreased in HA+" = "#CD4F39", "Nonsignificant" = "darkgray")
  ) +
  theme_bw(base_size = 14) + # change overall theme+
  theme(
    legend.position = "right",
    axis.text.y = y.axis.text,
    axis.text.x = x.axis.text,
    axis.title = axis.title,
    plot.title = plot.title.text,
    legend.text = legend.text,
    legend.title = legend.title.text,
    legend.key.size = unit(0.5, "cm"), # change legend key size
    legend.key.height = unit(0.5, "cm"), # change legend key height
    legend.key.width = unit(0.5, "cm")
  ) + # change legend key width
  xlab(expression(log[2]("Fold Change"))) + # Change X-Axis label
  ylab(expression(-log[10]("adjusted p-value"))) + # Change Y-Axis label
  geom_hline(yintercept = 1.3, colour = "darkgrey") + # Add p-adj value cutoff line
  scale_y_continuous(trans = "log1p") + # Scale y-axis due to large p-values
  geom_text_repel(aes(lfc,pval), 
                  label = ifelse(data$genelabels == TRUE, 
                  as.character(data$gene),""), 
                  box.padding = unit(1, "lines"),
                  hjust= 0.30, 
                  max.overlaps = Inf)
vol_final

#pdf(paste0(path, "output/Figure1D_VolcanoPlot.pdf"), width = 10, height = 8)
#vol_final
#dev.off()
```


# Gene Ontology Analysis {.tabset}

To generate GO analysis we first created a background of genes using any
genes that met the original DESEQ2 filtering criteria of \> 50 genes
counts per row and having a mappable Entrez ID. This generates the
necessary background to ensure that we are not just detecting background
functional pathways as we are utilizing this specific cell type. We then
filter our genes that have L2FC \< 0 or L2FC \> 0 to determine
directionality of the affected GO pathways. GO enrich is then used to
conduct the pathway analysis utilizing the Entrez IDs as inputs and an
"FDR" multiple testing correction.

### Pull only the Enriched samples from either side (\<-1 or \>1 LFC)

```{r}
# Generate Background for KEGG/GO
length(dds)
length(res1$entrez) 

background_mm10 <- subset(res1, is.na(entrez) == FALSE)
gene_matrix_background <- unique(sort(background_mm10$entrez))

# Remove any genes that do not have any entrez identifiers
results_sig_entrez <- subset(results_sig1, is.na(entrez) == FALSE)

# Create a matrix of gene log2 fold changes
gene_matrix <- results_sig_entrez$log2FoldChange

# Add the entrezID's as names for each logFC entry
names(gene_matrix) <- results_sig_entrez$entrez

# Pull only the Enriched samples from either side +/- 1 LFC
gene_matrix_enrich <- gene_matrix[which(gene_matrix > 1)]
gene_matrix_enrich2 <- gene_matrix[which(gene_matrix < -1)]

go_enrich1 <- clusterProfiler::enrichGO(gene = names(gene_matrix_enrich),
                      OrgDb = 'org.Mm.eg.db', 
                      readable = T,
                      universe = gene_matrix_background,
                      ont = "ALL",
                      pAdjustMethod = "fdr",
                      pvalueCutoff = 0.01, 
                      qvalueCutoff = 0.05)

go_enrich2 <- clusterProfiler::enrichGO(gene = names(gene_matrix_enrich2),
                      OrgDb = 'org.Mm.eg.db', 
                      readable = T,
                      universe = gene_matrix_background,
                      ont = "ALL",
                      pAdjustMethod = "fdr",
                      pvalueCutoff = 0.01, 
                      qvalueCutoff = 0.05)

KEGG_enrich1 <- clusterProfiler::enrichKEGG(gene = names(gene_matrix_enrich),
                      organism = 'mmu',
                      keyType = "ncbi-geneid",
                      pvalueCutoff = 0.01,
                      pAdjustMethod = "fdr",
                      universe = gene_matrix_background,
                      qvalueCutoff = 0.05)

KEGG_enrich2 <- clusterProfiler::enrichKEGG(gene = names(gene_matrix_enrich2),
                      organism = 'mmu',
                      keyType = "ncbi-geneid",
                      pvalueCutoff = 0.01,
                      pAdjustMethod = "fdr",
                      universe = gene_matrix_background,
                      qvalueCutoff = 0.05)

go_enrich_dataset1 <- as.data.frame(go_enrich1)
go_enrich_dataset2 <- as.data.frame(go_enrich2)

KEGG_enrich_dataset1 <- as.data.frame(KEGG_enrich1)
KEGG_enrich_dataset2 <- as.data.frame(KEGG_enrich2)

go_enrich_dataset1$Dataset <- "positive vs negative"
go_enrich_dataset2$Dataset <- "negative vs positive"

go_enrich_dataset1$Celltype <- "positive"
go_enrich_dataset2$Celltype <- "negative"

go_enrich_dataset1$Comparison <- "negative"
go_enrich_dataset2$Comparison <- "positive"

MergedGO1 <- merge(go_enrich_dataset1, go_enrich_dataset2, all=TRUE)

# Turn your 'Dataset' column into a character vector
MergedGO1$Dataset <- as.character(MergedGO1$Dataset)
# Then turn it back into a factor with the levels in the correct order
MergedGO1$Dataset <- factor(MergedGO1$Dataset, levels = c("positive vs negative", "negative vs positive"))

MergedGO1$Description_title_case <- stringr::str_to_title(MergedGO1$Description)
MergedGO1$GOPathways <- stringr::str_to_title(MergedGO1$Description_title_case)
MergedGO1$GOPathways <- stringr::str_wrap(MergedGO1$GOPathways, width = 10)

#S1 <- ggplot(MergedGO1, aes(x = Dataset, y = Description, size = Count, color = qvalue, group = Dataset)) +
#  geom_point(alpha = 1) +
#  theme_linedraw() + # change legend key width
#  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15)) +
#  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 50)) +
#  ggtitle("Enriched Pathways in PBN Nuclei Cell Types")
  #facet_grid(~Region+Comparison2, scales = "free_x")

#S2 <- S1 + scale_color_gradient(low = "red2", high = "mediumblue", space = "Lab") + scale_size(range = c(5, 10))
#S2
```

#### HA Positive Only

```{r}
## Only CGRP (positive)
GO1 <- go_enrich_dataset1

# Turn your 'Dataset' column into a character vector
GO1$Dataset <- as.character(GO1$Dataset)
# Then turn it back into a factor with the levels in the correct order
GO1$Dataset <- factor(GO1$Dataset, levels = c("positive vs negative"))

GO1$Description_title_case <- stringr::str_to_title(GO1$Description)
GO1$GOPathways <- stringr::str_to_title(GO1$Description_title_case)
GO1$GOPathways <- stringr::str_wrap(GO1$GOPathways, width = 10)

S1 <- ggplot(GO1, aes(x = Dataset, y = Description, size = Count, color = qvalue, group = Dataset)) +
  geom_point(alpha = 1) +
  theme_linedraw() + # change legend key width
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15)) +
  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 50)) +
  ggtitle("Enriched Pathways in PBN Nuclei Cell Types")
  #facet_grid(~Region+Comparison2, scales = "free_x")

S2 <- S1 + scale_color_gradient(low = "red2", high = "mediumblue", space = "Lab") + scale_size(range = c(1, 5))
S2
```

#### HA Negative Only

```{r fig.height=25, fig.width=20, message=FALSE, warning=FALSE}
## Only DAPI (negative)
GO2 <- go_enrich_dataset2

# Turn your 'Dataset' column into a character vector
GO2$Dataset <- as.character(GO2$Dataset)
# Then turn it back into a factor with the levels in the correct order
GO2$Dataset <- factor(GO2$Dataset, levels = c("negative vs positive"))

GO2$Description_title_case <- stringr::str_to_title(GO2$Description)
GO2$GOPathways <- stringr::str_to_title(GO2$Description_title_case)
GO2$GOPathways <- stringr::str_wrap(GO2$GOPathways, width = 10)

S1 <- ggplot(GO2, aes(x = Dataset, y = Description, size = Count, color = qvalue, group = Dataset)) +
  geom_point(alpha = 1) +
  theme_linedraw() + # change legend key width
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 5000)) +
  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 5000)) +
  ggtitle("Enriched Pathways in PBN Nuclei Cell Types")
  #facet_grid(~Region+Comparison2, scales = "free_x")

S2 <- S1 + scale_color_gradient(low = "red2", high = "mediumblue", space = "Lab") + scale_size(range = c(1, 5))
S2
```

### Set the Fold Enrichment for most important pathways

```{r}

go_filtered1 <- go_enrich_dataset1 %>%
  filter(FoldEnrichment >= 3)

go_filtered2 <- go_enrich_dataset2 %>%
  filter(FoldEnrichment >= 3)

go_filtered1$Dataset <- "positive vs negative"
go_filtered2$Dataset <- "negative vs positive"

go_filtered1$Celltype <- "positive"
go_filtered2$Celltype <- "negative"

go_filtered1$Comparison <- "negative"
go_filtered2$Comparison <- "positive"

```

#### HA Positive Only

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
GO1_filtered <- go_filtered1

# Turn your 'Dataset' column into a character vector
GO1_filtered$Dataset <- as.character(GO1_filtered$Dataset)
# Then turn it back into a factor with the levels in the correct order
GO1_filtered$Dataset <- factor(GO1_filtered$Dataset, levels = c("positive vs negative"))

GO1_filtered$Description_title_case <- stringr::str_to_title(GO1_filtered$Description)
GO1_filtered$GOPathways <- stringr::str_to_title(GO1_filtered$Description_title_case)
GO1_filtered$GOPathways <- stringr::str_wrap(GO1_filtered$GOPathways, width = 10)

S1 <- ggplot(GO1_filtered, aes(x = Count, y = reorder(Description, Count), fill = qvalue)) +
  geom_col(alpha = 1) +
  facet_wrap(~Dataset, scales = "free_y") +
  theme_linedraw() + # change legend key width
  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 50)) +
  ggtitle("Figure 1E. Enriched Pathways in HA+ Samples")

S2 <- S1 + scale_color_gradient(low = "red2", high = "mediumblue", space = "Lab") + scale_size(range = c(5, 10))
S2

#pdf(paste0(path, "output/Figure1E_GOHA.pdf"), width = 10, height = 8)
#S2
#dev.off()
```

#### HA Negative Only

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
GO2_filtered <- go_filtered2

# Turn your 'Dataset' column into a character vector
GO2_filtered$Dataset <- as.character(GO2_filtered$Dataset)
# Then turn it back into a factor with the levels in the correct order
GO2_filtered$Dataset <- factor(GO2_filtered$Dataset, levels = c("negative vs positive"))

GO2_filtered$Description_title_case <- stringr::str_to_title(GO2_filtered$Description)
GO2_filtered$GOPathways <- stringr::str_to_title(GO2_filtered$Description_title_case)
GO2_filtered$GOPathways <- stringr::str_wrap(GO2_filtered$GOPathways, width = 10)

S1 <- ggplot(GO2_filtered, aes(x = Count, y = reorder(Description, Count), fill = qvalue)) +
  geom_col(alpha = 1) +
  facet_wrap(~Dataset, scales = "free_y") +
  theme_linedraw() + # change legend key width
  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 50)) +
  ggtitle("Enriched Pathways in HA- Samples")

S2 <- S1 + scale_color_gradient(low = "red2", high = "mediumblue", space = "Lab") + scale_size(range = c(5, 10))
S2

#pdf(paste0(path, "output/Figure1F_GOHA.pdf"), width = 10, height = 8)
#S2
#dev.off()
```


```{r}

sessionInfo()

```
